---
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(printr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(pwr)
library("ggpubr")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_meta <- read.csv("~/master/wolf/dataflow/00-meta/wetzels2018_wu2018_fastq.csv")
colnames(df_meta)[1] <- "sra_accession"
df_meta$sra_accession <- as.character(df_meta$sra_accession)
df_meta$type <- as.character(df_meta$type)
df_meta$study <- as.character(df_meta$study)
df_meta$region <- as.character(df_meta$region)
df_meta$sample <- as.character(df_meta$sample)
df_meta$pack <- as.character(df_meta$pack)
df_meta$sibling <- as.character(df_meta$sibling)
df_meta$replicate <- as.character(df_meta$replicate)
df_meta$sample_id <- as.character(df_meta$sample_id)

df_classification  <- read.csv("~/master/wolf/dataflow/03-asv-taxonomy/wetzels2018-8_270_8_220-97-rdp.csv")
df_classification[,1] <- NULL

df <- read.delim("~/master/wolf/dataflow/03-asv-table/wetzels2018-8_270_8_220-97.txt", skip = 1, header = TRUE)
df$clustering <- 97

colnames(df)[1] <- 'asv_id'

col_to_gather <- names(df)[!(startsWith(names(df), "SRR") |  startsWith(names(df), "W"))]
df <- melt(df, id = col_to_gather)
colnames(df)[3:4] <- c('sra_accession', 'count')

df$asv_id <- as.character(df$asv_id)


df_normalization <- df %>%
  filter(clustering == 97) %>%
  group_by(sra_accession) %>%
  mutate(total_16s_reads = sum(count)) %>%
  ungroup() %>%
  select(sra_accession, total_16s_reads) %>%
  ungroup() %>%
  distinct() 

df_classification <- df_classification %>%
  select(asv_id, phylum, family, genus)


df_complete <- inner_join(df, df_meta) %>%
  inner_join(df_normalization) %>%
  inner_join(df_classification)

df_normalization_plot <- inner_join(df_normalization, df_meta)
```

## summary

- The Wolf skin and fecal samples (Wolf Science Centre - WSC) we're processed with dada2.
- All the data, after processing, was clustered together at 97% identity and taxonomically classified using RDB.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_metrics <- df_complete %>% 

  mutate(count_norm = (count / total_16s_reads)*100) %>%
  
  group_by(sample_id, asv_id) %>%
  
  mutate(count_norm = median(count_norm)) %>%
  
  ungroup() %>%
  
  unite(classification, c("phylum", "family", "genus"), sep = ";", remove = FALSE) %>%
  
  select(-sra_accession, -replicate, -total_16s_reads, -count) %>%
  distinct() 

df_metrics_trim <- df_metrics %>%
  filter(count_norm != 0) %>%
  group_by(asv_id) %>%
  mutate(nsamples = length(unique(sample_id))) %>%
  ungroup() %>%
  filter(nsamples > 2) %>%
  select(asv_id) %>%
  distinct()

df_metrics <- inner_join(df_metrics, df_metrics_trim)
```

\newpage

### phylum level grouping

- Skin and Fecal samples are very distinct at the phylum level. 

\vspace{12pt}

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
df_metrics_phylum <- df_metrics %>%
  group_by(sample_id, phylum) %>%
  mutate(phylum_percent = sum(count_norm)) %>%
  ungroup() %>%
  select(sample_id, phylum, phylum_percent, sample) %>%
  distinct() 

ggplot() + geom_bar(
  aes(y = phylum_percent, x = sample_id, fill = phylum),
  data = df_metrics_phylum,
  stat = "identity" ,
  position = "fill"
)  +labs(x="Category" , y= "Percentage (%)")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 16),
        strip.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 16), 
        axis.title.y = element_text(size = 20)) + 
     facet_wrap(.~sample, scales="free")+
  ylab("Percentage of Sample") +
  xlab("Individual Sample")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_phlycounts_counts <- df_metrics %>%
  select(asv_id, sample_id, count_norm) %>%
  spread(sample_id, count_norm)

df_phlycounts_counts <- df_phlycounts_counts[order(df_phlycounts_counts$asv_id),] 

otumat <- as.matrix(df_phlycounts_counts[,2:33])
rownames(otumat) <- as.data.frame(df_phlycounts_counts)[,1]

df_phlycounts_tax <- df_metrics %>%
  select(asv_id, phylum, family, genus) %>%
  unique()

df_phlycounts_tax$phylum <- as.character(df_phlycounts_tax$phylum)
df_phlycounts_tax$family <- as.character(df_phlycounts_tax$family)
df_phlycounts_tax$genus <- as.character(df_phlycounts_tax$genus)

df_phlycounts_tax[is.na(df_phlycounts_tax)] <- "NotAssigned"

df_phlycounts_tax <- df_phlycounts_tax[order(df_phlycounts_tax$asv_id),] 

taxmat <- as.matrix(df_phlycounts_tax[,2:4])
rownames(taxmat) <- as.data.frame(df_phlycounts_tax)[,1]

df_phylocounts_meta <- df_metrics %>% 
  select(sample_id, sample, pack, sibling) %>%
  unique() %>%
  rowwise() %>%
  ungroup()

df_phylocounts_meta <- as.data.frame(df_phylocounts_meta)

row.names(df_phylocounts_meta) <- df_phylocounts_meta[,1]
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=8}
library("phyloseq")
library("ape")
library(plyr)
OTU = otu_table(otumat, taxa_are_rows = TRUE)
TAX = tax_table(taxmat)
physeq = phyloseq(OTU, TAX)
fast_tree = read.tree("~/master/wolf/dataflow/02-trees/wetzels2018-8_270_8_220-97.out")
#random_tree = rtree(ntaxa(physeq), rooted=TRUE, tip.label=taxa_names(physeq))
samplesdata <- sample_data(df_phylocounts_meta)
physeq = phyloseq(OTU, TAX, samplesdata, fast_tree)
```

This is quite obvious is I make a network, where W4f calls between the WSC samples and the China samples.  

\vspace{12pt}

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=8}
net <- make_network(physeq, "samples", max.dist=0.8)
plot_network(net, physeq , color="pack", shape="sample", line_weight=0.3, label=NULL)
```

- We can also build all the different types of ordinations. Here, I will make the shape the region, as the two different regions in China seems to be more similar, and colour by the study. In the second version of the plot, I will colour by animals, as the dogs to stand out in a few ordinations. 

\vspace{12pt}


- I think the conclusion here is that the samples from China and the WSC are very different, with a single wolf from the WSC that appears to cluster quite a bit closer to the Chinese dogs and wolfs. 

### only looking at WSC feces data

- I will generate the ordinations again, with the first plots coloured by pack and the second coloured by sibling groups.
- Maybe a few of the packs a clustering a bit? For example, in the NMDS, 3 packs may cluster together - but nothing clear. 


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
dist = "bray"
ord_meths = c("DCA", "CCA", "RDA", "DPCoA", "NMDS", "MDS", "PCoA")
plist = llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="pack")
}, physeq, dist)
names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"

pdataframe <- subset(pdataframe, method != "NMDS")

p = ggplot(pdataframe, aes(Axis_1, Axis_2, color=pack, fill=pack, shape=sample))
p = p + geom_point(size=4) #+ geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + scale_colour_brewer(type="qual", palette="Set1")
p
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("doParallel")
library("foreach")

unifracmaxtrix <- UniFrac(physeq, weighted=TRUE, normalized=TRUE, parallel=FALSE, fast=TRUE)
df_unifrac <- as.data.frame(as.matrix(unifracmaxtrix))

library(gmodels)
library(tidyverse)


df_unifrac$id1 <- row.names(df_unifrac)

df_unifrac <- melt(df_unifrac, by = "id1")
colnames(df_unifrac)[2:3] <- c("id2", "UniFrac")

df_meta_trim1 <- df_meta %>%
  filter(study == "wetzels2018") %>%
  select(sample_id, sample, pack, sibling)

colnames(df_meta_trim1) <- c("id1", "sample1", "pack1", "sibling1") 

df_meta_trim2 <- df_meta %>%
  filter(study == "wetzels2018") %>%
  select(sample_id, sample, pack, sibling)

colnames(df_meta_trim2) <- c("id2", "sample2", "pack2", "sibling2") 

df_unifrac <- df_unifrac %>%
  filter(UniFrac != 0) %>%
  inner_join(df_meta_trim1) %>%
  inner_join(df_meta_trim2)

df_unifrac_test <- df_unifrac %>%
  
  rowwise() %>%
  mutate(same_pack = ifelse(pack1 == pack2, "SamePack", "DifferentPack")) %>%
  ungroup() %>%
  
  rowwise() %>%
  mutate(same_sample = ifelse(sample1 == sample2, "SameSample", "DifferentSample")) %>%
  ungroup() %>%
  
  group_by(same_sample) %>%
  
  mutate(mean_sample_unifrac = mean(UniFrac)) %>%
  mutate(ci_sample_unifrac = ci(UniFrac)) %>%
  
  ungroup()


```
```

