---
title: "Reciprocal Blast Hit Analysis"
author: "Anna Mueller"
date: '2018-11-09'
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(stringi)
<<<<<<< HEAD
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
blastdir <- '../dataflow/02-blast/'
files <- list.files(blastdir)
=======
library(ggplot2)
library(ggdendro)
library(reshape2)
library(grid)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
blastdir <- '~/master/rumen/dataflow/02-blast/'
files <- list.files(blastdir, pattern = "\\.txt$")
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
dflist <- list()

i <- 1
for (file in files){
  df <- read.table(paste(blastdir, file, sep = ''))
  df$file <- file
  df$V1 <- as.character(df$V1)
  df$V2 <- as.character(df$V2)
  dflist[[i]] <- df
  i <- i+1
}

df_compiled <- bind_rows(dflist)
colnames(df_compiled)[1:13] <- c("qseqid", "sseqid", "pident", "sstart", "send", "qstart", "qend", "evalue", "bitscore", "score", "qlen", "length", "sseq")

<<<<<<< HEAD
df_compiled$file <- gsub(".txt", "", df_compiled$file)



df_compiled <- df_compiled %>%
  separate(file, into = c("file1", "file2"), sep = "\\.") 
=======
df_compiled$file <- gsub("submission.assembly", "", df_compiled$file)
df_compiled$file <- gsub("\\.1", "1", df_compiled$file)
df_compiled$file <- gsub("\\.2", "2", df_compiled$file)
df_compiled$file <- gsub("\\.0", "0", df_compiled$file)
df_compiled$file <- gsub("_rename", "", df_compiled$file)
df_compiled$file <- gsub("_genomic", "", df_compiled$file)
df_compiled$file <- gsub("-", "", df_compiled$file)
df_compiled$file <- gsub("\\.txt", "", df_compiled$file)

df_compiled <- df_compiled %>%
  separate(file, into = c("file1", "file2"), sep = "\\.") %>%
  filter(length > 100)
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4

df_compiled$sseq <- NULL
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
df_forward <- df_compiled %>%
  select(qseqid, sseqid,file1, file2, pident) %>%
  rename(forward_pi = pident) %>%
  distinct()

df_reverse <- df_compiled %>%
  select(qseqid, sseqid, file1, file2, pident) %>%
  rename(reverse_pi = pident) %>%
  distinct()

colnames(df_reverse)[1:4] <- c("sseqid", "qseqid", "file2", "file1")

df_rbh <- inner_join(df_forward, df_reverse) %>%
  mutate(mean_pi = (forward_pi + reverse_pi)/2) %>%
  select(-forward_pi, -reverse_pi) %>%
  distinct() %>%
  rowwise() %>%
  mutate(samefile = ifelse(file1 == file2, "yes", "no")) %>%
  filter(samefile != "yes") %>%
  select(-samefile) %>%
  filter(mean_pi > 40)
```

<<<<<<< HEAD
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# header_dir <- '../dataflow/02-headers/'
# files <- list.files(header_dir)
# dflist <- list()
# 
# i <- 1
# for (file in files){
#   df <- read.csv(paste(header_dir, file, sep = ''), header = TRUE)
#   dflist[[i]] <- df
#   i <- i+1
# }
# 
# df_headers<- bind_rows(dflist)
# colnames(df_headers) <- c("qseqid", "header", "file1")
# 
# df_headers$file1 <- gsub(".fasta", "", df_headers$file1)
# df_headers$file1 <- gsub("_rename", "", df_headers$file1)
# 
 df_complete <- df_rbh #%>%
   #full_join(df_headers)

df_complete$sseqid[is.na(df_complete$sseqid)] <- "None"
df_complete$file2[is.na(df_complete$file2)] <- "None"
df_complete$mean_pi[is.na(df_complete$mean_pi)] <- 0.0

df_complete <- df_complete %>%
  separate(qseqid, into = c("genome",  "contig", "gene"), remove = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
=======
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=15, fig.height=10}
df_complete <- df_rbh 

df_complete$sseqid[is.na(df_complete$sseqid)] <- "None"
df_complete$file2[is.na(df_complete$file2)] <- "None"
df_complete$mean_pi[is.na(df_complete$mean_pi)] <- 0.0

#df_complete <- df_complete %>%
#  separate(qseqid, into = c("genome",  "contig", "gene"), remove = FALSE)

df_pi_shared <- df_rbh %>%
  select(file1, file2, mean_pi) %>%
  group_by(file1, file2) %>%
  mutate(mean_mean_pi = mean(mean_pi)) %>%
  ungroup() %>%
  select(-mean_pi) %>%
  distinct()


df_pi_shared_spread <- df_pi_shared %>%
  spread(file2, mean_mean_pi)

df_pi_shared_spread[is.na(df_pi_shared_spread)] <- 100

df_pi_shared_spread_scaled <- df_pi_shared_spread
df_pi_shared_spread[,2:12] <- scale(df_pi_shared_spread[,2:12])

df_pi_shared_spread.matrix <- as.matrix(df_pi_shared_spread[, -c(1)])
rownames(df_pi_shared_spread.matrix) <- df_pi_shared_spread$file1
df_pi_shared_spread.matrix.dendro <- as.dendrogram(hclust(d = dist(x = df_pi_shared_spread.matrix)))

dendro.plot <- ggdendrogram(data = df_pi_shared_spread.matrix.dendro, rotate = TRUE) 



df_pi_shared [is.na(df_pi_shared )] <- 100





order <- order.dendrogram(df_pi_shared_spread.matrix.dendro)

df_pi_shared$file1 <- factor(df_pi_shared$file1,
                               levels = df_pi_shared_spread_scaled$file1[order], 
                               ordered = TRUE)

heatmap.plot <- ggplot(data = df_pi_shared, aes(x = file2, y = file1)) +
  geom_tile(aes(fill = mean_mean_pi)) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  
        axis.text.y = element_text(size = 12)) +
  scale_fill_gradient2() +
  theme(legend.position = "top")


dendro.plot 
heatmap.plot

#grid.newpage()
#print(heatmap.plot, vp = viewport(x = 0.4, y = 0.5, width = 0.8, height = 1.0))
#print(dendro.plot, vp = viewport(x = 0.90, y = 0.445, width = 0.2, height = 1.0))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# rumen_group <- c("4300076", "4300142", "GCF_0000259251_ASM2592v1")
# vagina_group <- c("GCF_0007628651_04_NF40_HMP671v01", "GCF_0015465652_ASM154656v2")
# pos_group <- c("GCF_0001429651_ASM14296v1", "GCF_0002430151_Prev_macu_OT_289_V1", "GCF_0006135051_ASM61350v1", "GCF_0018146851_ASM181468v1", "GCF_0018148551_ASM181485v1", "GCF_0018153151_ASM181531v1")
# 
# df_list <- list()
# j <- 1
# 
# for (orf in unique(df_complete$qseqid)){
#   orf_df <- df_complete %>%
#   filter(qseqid == orf)
# 
#   orf_df$rumen_group <- sum(orf_df$file2 %in% rumen_group)
#   orf_df$vagina_group <- sum(orf_df$file2 %in% vagina_group)
#   orf_df$pos_group <- sum(orf_df$file2 %in% pos_group)
# 
#   df_list[[j]] <- orf_df
#   j <- j + 1
# }
# 
# df_complete_group_counts <- bind_rows(df_list)
# 
# write.csv(df_complete_group_counts, "~/Desktop/df_complete_group_counts.csv")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
df_complete_group_counts  <- read.csv("~/master/rumen/dataflow/03-blast-tables/df_complete_group_counts.csv") %>%
  select(-genome, -contig, -gene)

# these should be genes from one of the rumen genomes that is in all 5 non positive genomes

df_complete_group_counts_grouping <- df_complete_group_counts %>%
  mutate(non_pos = ifelse(pos_group == 0 & rumen_group == 2 & vagina_group == 2, "y", "n")) 


df_non_pos <- df_complete_group_counts_grouping %>%
  filter(non_pos == "y") %>%
  filter(file1 == "GCF_0000259251_ASM2592v1" | file1 == "4300076")

seqs <- as.data.frame(unique(df_non_pos$qseqid))
  
colnames(seqs) <- "prot_id"

write.csv(seqs, "~/master/rumen/dataflow/00-meta/non_positive_group.csv")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
df_complete <- df_complete %>%
  filter(file1 == "GCF_0000259251_ASM2592v1" | file1 == "4300076" | file1 == "4300142" | file1 == "GCF_0007628651_04_NF40_HMP671v01") %>%
  filter(file2 == "GCF_0000259251_ASM2592v1" | file2 == "4300076" | file2 == "4300142" | file2 == "GCF_0007628651_04_NF40_HMP671v01") 


df_complete$qseqid <- gsub("-submission.assembly", "", df_complete$qseqid)
df_complete$qseqid <- gsub("_ASM2592v1_genomic", "", df_complete$qseqid)

>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
genomes <- unique(df_complete$file1)
plots <- list()
i <- 1

for (genome1 in genomes){
  
    df_plot <- df_complete %>%
      filter(file1 == genome1)
    
    
    xbreaks <- df_plot$qseqid[seq(1, length(df_plot$qseqid), 100)]
    
    plot <- ggplot(df_plot, aes(x=qseqid)) +
    
      theme_few() +
      
<<<<<<< HEAD
      geom_point(aes(y = mean_pi, colour = file2), stat = "identity", size = 2) +
=======
      geom_point(aes(y = mean_pi, colour = file2), stat = "identity", size = 5) +
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
      
      theme(plot.title = element_text(size = 25), 
        axis.text.x = element_text(angle = 90, hjust = 1, size = 12),  
        axis.text.y = element_text(size = 20), 
        axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20)) + 
      
      ggtitle(paste(genome1)) + 
      
      ylab("Percent identity (%)") + 
      
      ylim(50,100) +
      
      #theme(legend.position="none") +
      
      theme(legend.text=element_text(size=20)) + 
      theme(legend.title=element_text(size=20)) +
      scale_x_discrete("Ordered genes",  breaks = xbreaks)
    
    plots[[i]] <- plot
    
    i <- i + 1
}
  
```

<<<<<<< HEAD
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=20, fig.height=4}
for (i in 1:length(plots)) {
  plot(plots[[i]])
  cat("\r\n\r\n")
}

#plot(plots[[1]])
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# genomes <- unique(df_rbh$file1)
# plots <- list()
# i <- 1
# 
# for (genome1 in genomes){
#   remaining_genomes <- genomes[!(genomes %in% genome1)]
#   for (genome2 in remaining_genomes){
#     df_plot <- df_rbh %>%
#       filter(file1 == genome1 & file2 == genome2)
#     
#     df_plot <- df_plot[order(df_plot$qseqid),]
#     df_plot$gene_num <- 1:nrow(df_plot) 
#     
#     plot <- ggplot(df_plot, aes(x=gene_num)) +
#     
#       theme_few() +
#       
#       geom_point(aes(y = mean_pi), stat = "identity", fill = "lightgrey", size = 2) +
#       
#       theme(plot.title = element_text(size = 25), 
#         axis.text.x = element_text(angle = 90, hjust = 1, size = 15),  
#         axis.text.y = element_text(size = 20), 
#         axis.title.x = element_text(size = 20), 
#         axis.title.y = element_text(size = 20)) + 
#       
#       ggtitle(paste(genome1, genome2, sep = "  v  ")) + 
#       
#       ylab("Percent identity (%)") + 
#     
#       xlab("Gene Number (order of genes)") + 
#       
#       theme(legend.text=element_text(size=20)) + 
#       theme(legend.title=element_text(size=20)) 
#     
#     plots[[i]] <- plot
#     
#     i <- i + 1
#   }
#   
#   
# }
```



=======
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=20, fig.height=5}
# for (i in 1:length(plots)) {
#   plot(plots[[i]])
#   cat("\r\n\r\n")
# }

plot(plots[[1]])
```
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
