---
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(printr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(pwr)
library("ggpubr")
library(stringi)
library(polycor)
library(gridExtra)
library(grid)
```

```{r global_options, R.options=knitr::opts_chunk$set(warning=FALSE, message=FALSE)}
```   

### Exploring patterns of Lactobacillus containing samples

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_complete <- read.csv("~/master/rumen/dataflow/03-asv-table/henderson2015_df_complete_temp.csv")
df_complete[,1] <- NULL

df_complete$asv_id <- as.character(df_complete$asv_id)
df_complete$GRCid <- as.character(df_complete$GRCid)
df_complete$sra_accession <- as.character(df_complete$sra_accession)
df_complete$LibraryName <- as.character(df_complete$LibraryName)

df_complete <- df_complete %>%
  rowwise() %>%
  mutate(replicate = stri_reverse(stri_split_fixed(stri_reverse(LibraryName),".",n = 2)[[1]][2])) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_metrics <- df_complete %>% 
  
  filter(domain == "Bacteria") %>%
  
  unique() %>%
  
  mutate(count_norm = (count / non.chimeric)*100) %>%
  
  group_by(asv_id, replicate) %>%
  mutate(n_count = length(asv_id)) %>%
  ungroup() %>%
  
<<<<<<< HEAD
  unite(classification, c("phylum", "family"), sep = ";", remove = FALSE) 
=======
  unite(classification, c("phylum", "family"), sep = ";", remove = FALSE) %>%
  filter(Simple.classification..animal. == "Cattle")
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4


df_metrics$classification <- gsub(";NA", "", df_metrics$classification)

df_samples <- df_metrics %>%
  select(GRCid, Starch.rich..diet., Concentrate.proportion..diet.) %>%
<<<<<<< HEAD
  distinct()
=======
  distinct() 
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_lacto_samples <- df_metrics %>%
  filter(genus == "Lactobacillus") %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName) %>%
<<<<<<< HEAD
  distinct()
=======
  distinct() 
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4


percent_lacto_samples_starch <- (table(df_lacto_samples$Starch.rich..diet.)[2]  / (table(df_lacto_samples$Starch.rich..diet.)[1] + table(df_lacto_samples$Starch.rich..diet.)[2])) * 100


percent_all_samples_starch <- (table(df_samples$Starch.rich..diet.)[2]  / (table(df_samples$Starch.rich..diet.)[1] + table(df_samples$Starch.rich..diet.)[2])) * 100
<<<<<<< HEAD
```

Where high starch samples make up `r percent_all_samples_starch` percent of the entire data set, if you subset the data set to only those samples where Lactobacillus is detected, then `r percent_lacto_samples_starch` are classified as high starch samples.  
=======


df_lacto_samples_amount <- df_lacto_samples %>%
  select(GRCid, med_count_norm) %>%
  distinct() %>%
  rename(lacto_count_norm = med_count_norm)
```

For this analysis, I am only looking at the cow samples.  
  
First thing I noticed: Where high starch samples make up `r percent_all_samples_starch` percent of the entire data set, if you subset the data set to only those samples where Lactobacillus is detected, then `r percent_lacto_samples_starch` are classified as high starch samples.  
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
  
I was then interested to subset the data to only high starch and then group then as with or without detectable Lactobacillus. Here, detectable Lactobacillus are being used for a marker of lower rumen heath.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
lacto_samples <- unique(df_lacto_samples$GRCid)

df_extreme_diet <- df_metrics %>%
  filter(Starch.rich..diet. == "y") %>%
  rowwise() %>%
  mutate(lacto_signal = ifelse(GRCid %in% lacto_samples, "pos", "neg")) %>%
  ungroup() %>%
  filter(Simple.classification..animal. == "Cattle")
<<<<<<< HEAD
```

Since all the data is subsetted to a similar diet, we can now look through each sequence or taxonomic group for difference between the Lactobacillus positive and Lactobacillus negative samples. This is asking the question which groups co-occur with Lactobacillus.

The dataset will be first be subsetted to only cows, and then those samples that have at least triplicates of the samples positive and negative for lactobacillus. Also I am focusing only on the Prevotellaceae as these are abundant, often implicated in SARA and have many groups in the data set. 
=======


n_lacto_samples <- length(lacto_samples)
n_extreme_lacto_samples  <- length(unique(df_extreme_diet$GRCid))
```

Since all the data is subsetted to a high starch diet, we can now look through each sequence or taxonomic group for difference between the Lactobacillus positive and Lactobacillus negative samples. 

The most abundant groups that seems to co-occur in the literature with the Lactobacillus are the Prevotellaceae. I therefore looked for sequences that were classified as Prevotellaceae where the mean abundance in either the Lactobacillus positive or negative samples is very high (over 1.5 %), and there was a significant difference between the 2 groups. This resulted in two selected sequences. For both sequences, there is an interesting trend, where specific data points jump up in adundance for only one of the groups (either Lactobacillus positive or negative).  

In doing the comparison between Prevotellaceae adbundant in samples with and without detectable Lactobacillus, it should be noted that there are `r n_lacto_samples` with detectable Lactobacillus and `r  n_extreme_lacto_samples - n_lacto_samples` without. 
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_extreme_diet <- df_extreme_diet %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(n_lacto_signal = length(count_norm)) %>%
  ungroup() %>%
  filter(n_lacto_signal > 3) %>%
  
  group_by(asv_id) %>%
  
  mutate(nlevels = length(unique(lacto_signal))) %>%
  
  ungroup() %>%
  
  filter(nlevels == 2) %>%
  
  filter(family != "NA")


df_extreme_diet_samples <- df_extreme_diet %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName, -med_count_norm, -classification, -domain, -phylum, -family, -genus, -n_lacto_signal) %>%
  distinct()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_ttest_list <- list()
i <- 1

for (group in unique(df_extreme_diet$asv_id)) {


  df_ttest<- df_extreme_diet %>%
    filter(asv_id == group)


      df_ttest$pval <- t.test(count_norm ~ lacto_signal, data = df_ttest)$p.value

      df_ttest_list[[i]] <- df_ttest
      i <- i + 1

}


df_ttest <- bind_rows(df_ttest_list) %>%
  distinct()
```

<<<<<<< HEAD
After doing t-tests between all of the sequences and plotting whose with differences, I noticed that the a few very abundant groups of Prevotellaceae positevely correlated with the detection of Lactobacillus, while many less abundant groups negatively correlate.
=======

>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4

\vspace{12pt}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_treatment_plot <- df_ttest  %>%
  filter(pval < 0.01) %>%
  arrange(pval) %>%
  filter(family == "Prevotellaceae") %>%
  distinct() %>%
  group_by(asv_id) %>%
  mutate(mean_seq = mean(count_norm)) %>%
  ungroup() %>%
<<<<<<< HEAD
  filter(mean_seq > 1)
=======
  
  group_by(asv_id, lacto_signal) %>%
  mutate(mean_seq_lacto = mean(count_norm)) %>%
  ungroup() 

df_abundance_select <- df_treatment_plot %>%
  filter(mean_seq_lacto > 2) %>%
  select(asv_id)
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4

df_direction <- df_treatment_plot %>%
  select(asv_id, lacto_signal, count_norm) %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(med_count_norm = median( count_norm)) %>%
  ungroup() %>%
  select(-count_norm) %>%
  distinct()

df_dir_list <- list()
i <- 1

for(seq in unique(df_direction$asv_id)){
  
  df_pos <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "pos")
  
  df_neg <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "neg")
  
  pos_med <- as.numeric(unique(df_pos$med_count_norm))
  neg_med <- as.numeric(unique(df_neg$med_count_norm))
  
  df_direction_seq <- df_direction %>%
    filter(asv_id == seq)
  
  if(pos_med > neg_med){
    
    df_direction_seq$direction <- "pos_cor"
    
  } else {
    df_direction_seq$direction <- "neg_cor"
  }
  
  df_dir_list[[i]] <- df_direction_seq
  i <- i + 1
  
}

df_direction <- bind_rows(df_dir_list) %>%
  select(asv_id, direction)

<<<<<<< HEAD
df_treatment_plot <- inner_join(df_treatment_plot, df_direction) 
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
=======
df_treatment_plot <- inner_join(df_treatment_plot, df_direction) %>%
  inner_join(df_abundance_select) %>% 
  left_join(df_lacto_samples_amount) %>%
  distinct()

df_treatment_plot$lacto_count_norm[is.na(df_treatment_plot$lacto_count_norm)] <- 0
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# I then wanted to know if the concentration of Lactobacillus correlated with the other variables in the data set, to understand what is causing or confounding the effect. 


df_multi_correlation <- df_treatment_plot  %>%
  select(Age.classification..animal., Gender..animal., Forage.proportion..diet., Concentrate.proportion..diet., Pectin.rich...diet., Weeks.on.diet..diet., Season..location., lacto_signal, Forage.concentrate.category..diet., lacto_count_norm) %>%
  distinct() %>%
  arrange(desc(lacto_count_norm))

df_multi_correlation$Concentrate.proportion..diet. <- as.numeric(df_multi_correlation$Concentrate.proportion..diet.)

df_multi_correlation <- as.data.frame(df_multi_correlation)
df_multi_correlation <- df_multi_correlation[complete.cases(df_multi_correlation), ]

df_multi_correlation$lacto_signal <- as.factor(df_multi_correlation$lacto_signal)

df_het_cor <- as.data.frame(hetcor(df_multi_correlation)$correlations)

df_het_cor_select <- df_het_cor %>%
  select(lacto_count_norm) %>%
  arrange(desc(lacto_count_norm))

row.names(df_het_cor_select) <- row.names(df_het_cor)


#df_het_cor_select
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}




>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
groups <- unique(df_treatment_plot$asv_id)
df_treatment_plot$lacto_signal <- factor(df_treatment_plot$lacto_signal, levels = c("neg", "pos"))

plots_neg <- list()
plots_pos <- list()

i <- 1
j <- 1

for (group in groups){
  
    df_p <- df_treatment_plot %>%
      filter(asv_id == group)
    
    taxa <- unique(df_p$classification)
    dir <- unique(df_p$direction)
# plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA, add = "dotplot", add.params = list(color = "Gender..animal.")) +
#           ylab("Normalized Count") +
#           xlab("Lactobacillus Signal") +
#           theme(strip.text = element_text(size = 15),
#           plot.title = element_text(size = 30),
#           axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
#           axis.text.y = element_text(size = 20),
#           axis.title.x = element_text(size = 20),
#           axis.title.y = element_text(size = 20)) + 
#         
# 
#           #ggtitle(taxa) +
#   
#           facet_grid(asv_id ~ ., scales = "free")


<<<<<<< HEAD
plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA) +
          ylab("Normalized Count") +
          xlab("Lactobacillus Signal") +
          theme(strip.text = element_text(size = 15),
=======
plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA, add="jitter") +
          ylab("Normalized Count") +
          xlab("Lactobacillus Signal") +
          theme(strip.text = element_text(size = 10),
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
          plot.title = element_text(size = 30),
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20)) + 
        

          #ggtitle(taxa) +
  
          facet_grid(asv_id ~ ., scales = "free")
    

if(dir == "neg_cor"){
plots_neg[[j]] <- plot
j <- j + 1
} else{
plots_pos[[i]] <- plot
i <- i + 1
}
}
```

<<<<<<< HEAD
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=7}
grid.arrange(grobs = plots_pos, ncol = 1, top=textGrob("Prevotellaceae co-occuring with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=25, fig.height=25}
grid.arrange(grobs = plots_neg, ncol = 1, top=textGrob("Prevotellaceae negatively correlated with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```

Now I need to check what of the other factors I have in my dataset correlate with the presence of lactobacillus. This is essentially to understand what variables could be confounding this effect.  
  
Appears that there is some correlation between the pectin rich diets, age and the Lactobacillus signal. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_multi_correlation <- df_treatment_plot %>%
  select(Age.classification..animal., Gender..animal., Forage.proportion..diet., Concentrate.proportion..diet., Pectin.rich...diet., Weeks.on.diet..diet., Season..location., lacto_signal) %>%
  distinct()

df_multi_correlation$Concentrate.proportion..diet. <- as.numeric(df_multi_correlation$Concentrate.proportion..diet.)

df_multi_correlation <- as.data.frame(df_multi_correlation)
df_multi_correlation <- df_multi_correlation[complete.cases(df_multi_correlation), ]



df_het_cor <- as.data.frame(hetcor(df_multi_correlation)$correlations)

df_het_cor
```
=======
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=15, fig.height=10}
grid.arrange(plots_pos[[1]], plots_neg[[1]], ncol = 2, top=textGrob("Prevotellaceae co-occuring with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```


>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_direction <- df_direction %>%
  inner_join(df_treatment_plot) %>%
  select(asv_id, direction) %>%
  distinct()

write.csv(df_direction,  "~/master/rumen/dataflow/00-meta/henderson2015_lactobacillus_association.csv")
<<<<<<< HEAD
```
=======
```
>>>>>>> bc63259588b9b18c1f659eccbc5b966d470663f4
