---
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(printr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(pwr)
library("ggpubr")
library(stringi)
library(polycor)
library(gridExtra)
library(grid)
```

```{r global_options, R.options=knitr::opts_chunk$set(warning=FALSE, message=FALSE)}
```   

### Exploring patterns of Lactobacillus containing samples

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_complete <- read.csv("~/master/rumen/dataflow/03-asv-table/henderson2015_df_complete_temp.csv")
df_complete[,1] <- NULL

df_complete$asv_id <- as.character(df_complete$asv_id)
df_complete$GRCid <- as.character(df_complete$GRCid)
df_complete$sra_accession <- as.character(df_complete$sra_accession)
df_complete$LibraryName <- as.character(df_complete$LibraryName)

df_complete <- df_complete %>%
  rowwise() %>%
  mutate(replicate = stri_reverse(stri_split_fixed(stri_reverse(LibraryName),".",n = 2)[[1]][2])) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_metrics <- df_complete %>% 
  
  filter(domain == "Bacteria") %>%
  
  unique() %>%
  
  mutate(count_norm = (count / non.chimeric)*100) %>%
  
  group_by(asv_id, replicate) %>%
  mutate(n_count = length(asv_id)) %>%
  ungroup() %>%
  
  unite(classification, c("phylum", "family"), sep = ";", remove = FALSE) %>%
  filter(Simple.classification..animal. == "Cattle")


df_metrics$classification <- gsub(";NA", "", df_metrics$classification)

df_samples <- df_metrics %>%
  select(GRCid, Starch.rich..diet., Concentrate.proportion..diet.) %>%
  distinct() 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_lacto_samples <- df_metrics %>%
  filter(genus == "Lactobacillus") %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName) %>%
  distinct() 


percent_lacto_samples_starch <- (table(df_lacto_samples$Starch.rich..diet.)[2]  / (table(df_lacto_samples$Starch.rich..diet.)[1] + table(df_lacto_samples$Starch.rich..diet.)[2])) * 100


percent_all_samples_starch <- (table(df_samples$Starch.rich..diet.)[2]  / (table(df_samples$Starch.rich..diet.)[1] + table(df_samples$Starch.rich..diet.)[2])) * 100


df_lacto_samples_amount <- df_lacto_samples %>%
  select(GRCid, med_count_norm) %>%
  distinct() %>%
  rename(lacto_count_norm = med_count_norm)
```

For this analysis, I am only looking at the cow samples.  
  
First thing I noticed: Where high starch samples make up `r percent_all_samples_starch` percent of the entire data set, if you subset the data set to only those samples where Lactobacillus is detected, then `r percent_lacto_samples_starch` are classified as high starch samples.  
  
I was then interested to subset the data to only high starch and then group then as with or without detectable Lactobacillus. Here, detectable Lactobacillus are being used for a marker of lower rumen heath.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
lacto_samples <- unique(df_lacto_samples$GRCid)

df_extreme_diet <- df_metrics %>%
  filter(Starch.rich..diet. == "y") %>%
  rowwise() %>%
  mutate(lacto_signal = ifelse(GRCid %in% lacto_samples, "pos", "neg")) %>%
  ungroup() %>%
  filter(Simple.classification..animal. == "Cattle")


n_lacto_samples <- length(lacto_samples)
n_extreme_lacto_samples  <- length(unique(df_extreme_diet$GRCid))
```

Since all the data is subsetted to a high starch diet, we can now look through each sequence or taxonomic group for difference between the Lactobacillus positive and Lactobacillus negative samples. 

The most abundant groups that seems to co-occur in the literature with the Lactobacillus are the Prevotellaceae. I therefore looked for sequences that were classified as Prevotellaceae where the mean abundance in either the Lactobacillus positive or negative samples is very high (over 1.5 %), and there was a significant difference between the 2 groups. This resulted in two selected sequences. For both sequences, there is an interesting trend, where specific data points jump up in adundance for only one of the groups (either Lactobacillus positive or negative).  

In doing the comparison between Prevotellaceae adbundant in samples with and without detectable Lactobacillus, it should be noted that there are `r n_lacto_samples` with detectable Lactobacillus and `r  n_extreme_lacto_samples - n_lacto_samples` without. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_extreme_diet <- df_extreme_diet %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(n_lacto_signal = length(count_norm)) %>%
  ungroup() %>%
  filter(n_lacto_signal > 3) %>%
  
  group_by(asv_id) %>%
  
  mutate(nlevels = length(unique(lacto_signal))) %>%
  
  ungroup() %>%
  
  filter(nlevels == 2) %>%
  
  filter(family != "NA")


df_extreme_diet_samples <- df_extreme_diet %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName, -med_count_norm, -classification, -domain, -phylum, -family, -genus, -n_lacto_signal) %>%
  distinct()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_ttest_list <- list()
i <- 1

for (group in unique(df_extreme_diet$asv_id)) {


  df_ttest<- df_extreme_diet %>%
    filter(asv_id == group)


      df_ttest$pval <- t.test(count_norm ~ lacto_signal, data = df_ttest)$p.value

      df_ttest_list[[i]] <- df_ttest
      i <- i + 1

}


df_ttest <- bind_rows(df_ttest_list) %>%
  distinct()
```



\vspace{12pt}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_treatment_plot <- df_ttest  %>%
  filter(pval < 0.01) %>%
  arrange(pval) %>%
  filter(family == "Prevotellaceae") %>%
  distinct() %>%
  group_by(asv_id) %>%
  mutate(mean_seq = mean(count_norm)) %>%
  ungroup() %>%
  
  group_by(asv_id, lacto_signal) %>%
  mutate(mean_seq_lacto = mean(count_norm)) %>%
  ungroup() 

df_abundance_select <- df_treatment_plot %>%
  filter(mean_seq_lacto > 2) %>%
  select(asv_id)

df_direction <- df_treatment_plot %>%
  select(asv_id, lacto_signal, count_norm) %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(med_count_norm = median( count_norm)) %>%
  ungroup() %>%
  select(-count_norm) %>%
  distinct()

df_dir_list <- list()
i <- 1

for(seq in unique(df_direction$asv_id)){
  
  df_pos <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "pos")
  
  df_neg <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "neg")
  
  pos_med <- as.numeric(unique(df_pos$med_count_norm))
  neg_med <- as.numeric(unique(df_neg$med_count_norm))
  
  df_direction_seq <- df_direction %>%
    filter(asv_id == seq)
  
  if(pos_med > neg_med){
    
    df_direction_seq$direction <- "pos_cor"
    
  } else {
    df_direction_seq$direction <- "neg_cor"
  }
  
  df_dir_list[[i]] <- df_direction_seq
  i <- i + 1
  
}

df_direction <- bind_rows(df_dir_list) %>%
  select(asv_id, direction)

df_treatment_plot <- inner_join(df_treatment_plot, df_direction) %>%
  inner_join(df_abundance_select) %>% 
  left_join(df_lacto_samples_amount) %>%
  distinct()

df_treatment_plot$lacto_count_norm[is.na(df_treatment_plot$lacto_count_norm)] <- 0
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# I then wanted to know if the concentration of Lactobacillus correlated with the other variables in the data set, to understand what is causing or confounding the effect. 


df_multi_correlation <- df_treatment_plot  %>%
  select(Age.classification..animal., Gender..animal., Forage.proportion..diet., Concentrate.proportion..diet., Pectin.rich...diet., Weeks.on.diet..diet., Season..location., lacto_signal, Forage.concentrate.category..diet., lacto_count_norm) %>%
  distinct() %>%
  arrange(desc(lacto_count_norm))

df_multi_correlation$Concentrate.proportion..diet. <- as.numeric(df_multi_correlation$Concentrate.proportion..diet.)

df_multi_correlation <- as.data.frame(df_multi_correlation)
df_multi_correlation <- df_multi_correlation[complete.cases(df_multi_correlation), ]

df_multi_correlation$lacto_signal <- as.factor(df_multi_correlation$lacto_signal)

df_het_cor <- as.data.frame(hetcor(df_multi_correlation)$correlations)

df_het_cor_select <- df_het_cor %>%
  select(lacto_count_norm) %>%
  arrange(desc(lacto_count_norm))

row.names(df_het_cor_select) <- row.names(df_het_cor)


#df_het_cor_select
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}




groups <- unique(df_treatment_plot$asv_id)
df_treatment_plot$lacto_signal <- factor(df_treatment_plot$lacto_signal, levels = c("neg", "pos"))

plots_neg <- list()
plots_pos <- list()

i <- 1
j <- 1

for (group in groups){
  
    df_p <- df_treatment_plot %>%
      filter(asv_id == group)
    
    taxa <- unique(df_p$classification)
    dir <- unique(df_p$direction)
# plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA, add = "dotplot", add.params = list(color = "Gender..animal.")) +
#           ylab("Normalized Count") +
#           xlab("Lactobacillus Signal") +
#           theme(strip.text = element_text(size = 15),
#           plot.title = element_text(size = 30),
#           axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
#           axis.text.y = element_text(size = 20),
#           axis.title.x = element_text(size = 20),
#           axis.title.y = element_text(size = 20)) + 
#         
# 
#           #ggtitle(taxa) +
#   
#           facet_grid(asv_id ~ ., scales = "free")


plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA, add="jitter") +
          ylab("Normalized Count") +
          xlab("Lactobacillus Signal") +
          theme(strip.text = element_text(size = 10),
          plot.title = element_text(size = 30),
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20)) + 
        

          #ggtitle(taxa) +
  
          facet_grid(asv_id ~ ., scales = "free")
    

if(dir == "neg_cor"){
plots_neg[[j]] <- plot
j <- j + 1
} else{
plots_pos[[i]] <- plot
i <- i + 1
}
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=15, fig.height=10}
grid.arrange(plots_pos[[1]], plots_neg[[1]], ncol = 2, top=textGrob("Prevotellaceae co-occuring with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_direction <- df_direction %>%
  inner_join(df_treatment_plot) %>%
  select(asv_id, direction) %>%
  distinct()

write.csv(df_direction,  "~/master/rumen/dataflow/00-meta/henderson2015_lactobacillus_association.csv")
```