---
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(printr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(pwr)
library("ggpubr")
library(stringi)
library(polycor)
library(gridExtra)
library(grid)
```

```{r global_options, R.options=knitr::opts_chunk$set(warning=FALSE, message=FALSE)}
```   

### Exploring patterns of Lactobacillus containing samples

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_complete <- read.csv("~/master/rumen/dataflow/03-asv-table/henderson2015_df_complete_temp.csv")
df_complete[,1] <- NULL

df_complete$asv_id <- as.character(df_complete$asv_id)
df_complete$GRCid <- as.character(df_complete$GRCid)
df_complete$sra_accession <- as.character(df_complete$sra_accession)
df_complete$LibraryName <- as.character(df_complete$LibraryName)

df_complete <- df_complete %>%
  rowwise() %>%
  mutate(replicate = stri_reverse(stri_split_fixed(stri_reverse(LibraryName),".",n = 2)[[1]][2])) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_metrics <- df_complete %>% 
  
  filter(domain == "Bacteria") %>%
  
  unique() %>%
  
  mutate(count_norm = (count / non.chimeric)*100) %>%
  
  group_by(asv_id, replicate) %>%
  mutate(n_count = length(asv_id)) %>%
  ungroup() %>%
  
  unite(classification, c("phylum", "family"), sep = ";", remove = FALSE) 


df_metrics$classification <- gsub(";NA", "", df_metrics$classification)

df_samples <- df_metrics %>%
  select(GRCid, Starch.rich..diet., Concentrate.proportion..diet.) %>%
  distinct()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_lacto_samples <- df_metrics %>%
  filter(genus == "Lactobacillus") %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName) %>%
  distinct()


percent_lacto_samples_starch <- (table(df_lacto_samples$Starch.rich..diet.)[2]  / (table(df_lacto_samples$Starch.rich..diet.)[1] + table(df_lacto_samples$Starch.rich..diet.)[2])) * 100


percent_all_samples_starch <- (table(df_samples$Starch.rich..diet.)[2]  / (table(df_samples$Starch.rich..diet.)[1] + table(df_samples$Starch.rich..diet.)[2])) * 100
```

Where high starch samples make up `r percent_all_samples_starch` percent of the entire data set, if you subset the data set to only those samples where Lactobacillus is detected, then `r percent_lacto_samples_starch` are classified as high starch samples.  
  
I was then interested to subset the data to only high starch and then group then as with or without detectable Lactobacillus. Here, detectable Lactobacillus are being used for a marker of lower rumen heath.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
lacto_samples <- unique(df_lacto_samples$GRCid)

df_extreme_diet <- df_metrics %>%
  filter(Starch.rich..diet. == "y") %>%
  rowwise() %>%
  mutate(lacto_signal = ifelse(GRCid %in% lacto_samples, "pos", "neg")) %>%
  ungroup() %>%
  filter(Simple.classification..animal. == "Cattle")
```

Since all the data is subsetted to a similar diet, we can now look through each sequence or taxonomic group for difference between the Lactobacillus positive and Lactobacillus negative samples. This is asking the question which groups co-occur with Lactobacillus.

The dataset will be first be subsetted to only cows, and then those samples that have at least triplicates of the samples positive and negative for lactobacillus. Also I am focusing only on the Prevotellaceae as these are abundant, often implicated in SARA and have many groups in the data set. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_extreme_diet <- df_extreme_diet %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(n_lacto_signal = length(count_norm)) %>%
  ungroup() %>%
  filter(n_lacto_signal > 3) %>%
  
  group_by(asv_id) %>%
  
  mutate(nlevels = length(unique(lacto_signal))) %>%
  
  ungroup() %>%
  
  filter(nlevels == 2) %>%
  
  filter(family != "NA")


df_extreme_diet_samples <- df_extreme_diet %>%
  
  group_by(GRCid) %>%
  mutate(med_count_norm = median(count_norm)) %>%
  ungroup() %>%
  
  select(-asv_id, -count, -count_norm, -LibraryName, -med_count_norm, -classification, -domain, -phylum, -family, -genus, -n_lacto_signal) %>%
  distinct()
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_ttest_list <- list()
i <- 1

for (group in unique(df_extreme_diet$asv_id)) {


  df_ttest<- df_extreme_diet %>%
    filter(asv_id == group)


      df_ttest$pval <- t.test(count_norm ~ lacto_signal, data = df_ttest)$p.value

      df_ttest_list[[i]] <- df_ttest
      i <- i + 1

}


df_ttest <- bind_rows(df_ttest_list) %>%
  distinct()
```

After doing t-tests between all of the sequences and plotting whose with differences, I noticed that the a few very abundant groups of Prevotellaceae positevely correlated with the detection of Lactobacillus, while many less abundant groups negatively correlate.

\vspace{12pt}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_treatment_plot <- df_ttest  %>%
  filter(pval < 0.01) %>%
  arrange(pval) %>%
  filter(family == "Prevotellaceae") %>%
  distinct() %>%
  group_by(asv_id) %>%
  mutate(mean_seq = mean(count_norm)) %>%
  ungroup() %>%
  filter(mean_seq > 1)

df_direction <- df_treatment_plot %>%
  select(asv_id, lacto_signal, count_norm) %>%
  group_by(asv_id, lacto_signal) %>%
  mutate(med_count_norm = median( count_norm)) %>%
  ungroup() %>%
  select(-count_norm) %>%
  distinct()

df_dir_list <- list()
i <- 1

for(seq in unique(df_direction$asv_id)){
  
  df_pos <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "pos")
  
  df_neg <- df_direction %>%
    filter(asv_id == seq) %>%
    filter(lacto_signal == "neg")
  
  pos_med <- as.numeric(unique(df_pos$med_count_norm))
  neg_med <- as.numeric(unique(df_neg$med_count_norm))
  
  df_direction_seq <- df_direction %>%
    filter(asv_id == seq)
  
  if(pos_med > neg_med){
    
    df_direction_seq$direction <- "pos_cor"
    
  } else {
    df_direction_seq$direction <- "neg_cor"
  }
  
  df_dir_list[[i]] <- df_direction_seq
  i <- i + 1
  
}

df_direction <- bind_rows(df_dir_list) %>%
  select(asv_id, direction)

df_treatment_plot <- inner_join(df_treatment_plot, df_direction) 
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
groups <- unique(df_treatment_plot$asv_id)
df_treatment_plot$lacto_signal <- factor(df_treatment_plot$lacto_signal, levels = c("neg", "pos"))

plots_neg <- list()
plots_pos <- list()

i <- 1
j <- 1

for (group in groups){
  
    df_p <- df_treatment_plot %>%
      filter(asv_id == group)
    
    taxa <- unique(df_p$classification)
    dir <- unique(df_p$direction)
# plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA, add = "dotplot", add.params = list(color = "Gender..animal.")) +
#           ylab("Normalized Count") +
#           xlab("Lactobacillus Signal") +
#           theme(strip.text = element_text(size = 15),
#           plot.title = element_text(size = 30),
#           axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
#           axis.text.y = element_text(size = 20),
#           axis.title.x = element_text(size = 20),
#           axis.title.y = element_text(size = 20)) + 
#         
# 
#           #ggtitle(taxa) +
#   
#           facet_grid(asv_id ~ ., scales = "free")


plot <- ggboxplot(df_p, x = "lacto_signal", y = "count_norm", outlier.shape = NA) +
          ylab("Normalized Count") +
          xlab("Lactobacillus Signal") +
          theme(strip.text = element_text(size = 15),
          plot.title = element_text(size = 30),
          axis.text.x = element_text(angle = 90, hjust = 1, size = 20),
          axis.text.y = element_text(size = 20),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20)) + 
        

          #ggtitle(taxa) +
  
          facet_grid(asv_id ~ ., scales = "free")
    

if(dir == "neg_cor"){
plots_neg[[j]] <- plot
j <- j + 1
} else{
plots_pos[[i]] <- plot
i <- i + 1
}
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=7}
grid.arrange(grobs = plots_pos, ncol = 1, top=textGrob("Prevotellaceae co-occuring with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=25, fig.height=25}
grid.arrange(grobs = plots_neg, ncol = 1, top=textGrob("Prevotellaceae negatively correlated with Lactobacillus",gp=gpar(fontsize=20,font=3)))
```

Now I need to check what of the other factors I have in my dataset correlate with the presence of lactobacillus. This is essentially to understand what variables could be confounding this effect.  
  
Appears that there is some correlation between the pectin rich diets, age and the Lactobacillus signal. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_multi_correlation <- df_treatment_plot %>%
  select(Age.classification..animal., Gender..animal., Forage.proportion..diet., Concentrate.proportion..diet., Pectin.rich...diet., Weeks.on.diet..diet., Season..location., lacto_signal) %>%
  distinct()

df_multi_correlation$Concentrate.proportion..diet. <- as.numeric(df_multi_correlation$Concentrate.proportion..diet.)

df_multi_correlation <- as.data.frame(df_multi_correlation)
df_multi_correlation <- df_multi_correlation[complete.cases(df_multi_correlation), ]



df_het_cor <- as.data.frame(hetcor(df_multi_correlation)$correlations)

df_het_cor
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df_direction <- df_direction %>%
  inner_join(df_treatment_plot) %>%
  select(asv_id, direction) %>%
  distinct()

write.csv(df_direction,  "~/master/rumen/dataflow/00-meta/henderson2015_lactobacillus_association.csv")
```
