---
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(printr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(pwr)
library("ggpubr")
library(stringi)
library(polycor)
library(gridExtra)
library(grid)
```

```{r global_options, R.options=knitr::opts_chunk$set(warning=FALSE, message=FALSE)}
```   

Two sequences were selected from the analysis henderson2015_lactobacillus_grouping.Rmd. Since these sequences were representatives from 97% indentity clusters, I blasted them against the 100% asvs and then pulled out all the seqs above 97% identity here. Then the sequences were given new headers.  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- read.delim("~/master/rumen/dataflow/03-blast-tables/lacto_prevo_100_mapped", header=FALSE)

colnames(df) <- c("qseqid", "sseqid", "pident", "sstart", "send", "qstart", "qend", "evalue", "bitscore", "score", "qlen", "length", "sseq")

df_select <- df %>%
  filter(length > 323) %>%
  filter(pident > 97)


df_select <- as.data.frame(df_select)
df_select$sseq <- as.character(df_select$sseq)
df_select$qseqid <- as.character(df_select$qseqid)

df_select_list <- list()
k <- 1

for (j in 1:nrow(df_select)) {
  
  seq_id <- df_select[j,1]
  
  if (seq_id == "cebd44fd058a519d1dc298117fae3289_negative_correlation" ){ 
      header <- paste(">", 'CEB_N.', toString(j), sep = "") }
   else {
      header <- paste(">", '8CB_P.', toString(j), sep = "") }

  
  df_select_list[[k]] <- header
  k <- k + 1
  
  seq <- df_select[j,13]
  
  df_select_list[[k]] <- seq
  k <- k + 1
  
  
  
}


selected_long <- as.matrix(unlist(rbind(df_select_list)))
selected_long <- gsub("\"", "", selected_long)

write.table(selected_long, "~/master/rumen/dataflow/01-nucl/lacto_prevo_100.fasta", row.names = FALSE, 
            col.names = FALSE, quote = FALSE)
```

The resulting file was blasted against the rumen genome database as well as all of the Prevotellaceae genomes from NCBI. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
df1 <- read.delim("~/master/rumen/dataflow/03-blast-tables/lacto_prevo_100_prevo_genomes_mapped", header=FALSE)
df1$database <- "Prevotella"

df2 <- read.delim("~/master/rumen/dataflow/03-blast-tables/lacto_prevo_100_rumen_genomes_mapped", header=FALSE)
df2$database <- "Rumen"


df <- bind_rows(df1, df2)

colnames(df) <- c("qseqid", "sseqid", "pident", "sstart", "send", "qstart", "qend", "evalue", "bitscore", "score", "qlen", "length", "sseq", "database")

df$length <- as.numeric(df$length)
df$pident <- as.numeric(df$pident)
df$bitscore <- as.numeric(df$bitscore)

df_select <- df %>%
  filter(length > 300) %>%
  filter(pident > 93) %>%
  rowwise() %>%
  mutate(seq_num = stri_reverse(stri_split_fixed(stri_reverse(sseqid),"_",n = 2)[[1]][1])) %>%
  mutate(genome = stri_reverse(stri_split_fixed(stri_reverse(sseqid),"_",n = 2)[[1]][2])) 

df_select$genome <- gsub("_genomic", "", df_select$genome)
df_select$genome <- gsub("-submission.assembly", "", df_select$genome)
df_select$genome <- gsub("_annotated_assembly", "", df_select$genome)





df_select <- as.data.frame(df_select)
df_select$sseq <- as.character(df_select$sseq)

df_select_list <- list()
k <- 1

for (j in 1:nrow(df_select)) {
  
  database <- df_select[j,14]
  genome <- df_select[j,16]
  header <- paste(">", database, "_", genome,  sep = "")

  
  df_select_list[[k]] <- header
  k <- k + 1
  
  seq <- df_select[j,13]
  
  df_select_list[[k]] <- seq
  k <- k + 1
  
  
  
}


selected_long <- as.matrix(unlist(rbind(df_select_list)))
selected_long <- gsub("\"", "", selected_long)

write.table(selected_long, "~/master/rumen/dataflow/01-nucl/lacto_prevo_100_genomes_blast.fasta", row.names = FALSE, 
            col.names = FALSE, quote = FALSE)


```